---
title: "da4-morning"
format: html
editor: visual
---
Add the libraries
```{r}
library(tidyverse)
```

Questions:
Why do you need a `seq_along`? dont the for explains that?

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for(i in quarter_splits) {
  half_time <- quarter_splits[i] + quarter_splits[i+1]
  print(half_time)
}

for(i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] + quarter_splits[i+1]
  print(half_time)
}

```


##Intro to Functions
### Make a funtion to dougle the user input

```{r}

double_this <- function(my_value) {
  print(2*my_value)
}

double_this(200)

double_this(c(1,2,100,125,175))
```


#### Write a function that asks a user `name` & `age`, and returns an exlamation, like "I´m [NAME], and I am [AGE] years old

```{r}

age <- function(name,age) {
  print(paste("I am,", name, "and I am", age, "years old"))
}

age(age= 29, name = "Javier")
age( "Javier" , 29)
```

### Make a function that has conditionals

```{r}

convert_animal_age <- function(x,y) {
  if (x == "dog") {
    print(paste("Your dog is", y*7, "human years old"))
  }
  else if (x == "goat") {
    print(paste("Your goat is", y*4.7, "human years old"))
  }
  else{
    print("We dont have the convertion data for that animal")
  }
}
convert_animal_age("ele",5)
```

Create a Data.Frame and play with function
```{r}

dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken",
)


dog_menu <- function(name) {
  my_subset <- filter(dog_choice, dog_name == name)
  print(paste("My name is", my_subset$dog_name, "and my favorite food is", my_subset$food))
}

dog_menu("Waffle")

```

### Function meets for loops

```{r}
#Apply a mean over specific columns within a data frames

means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    col_name <- colnames(df[i])
    print(paste("The mean value of", col_name, "is", col_mean))
  }
}

means_over_columns(mtcars)
```



### Interation in matrix

```{r}
x <- c(1,2,3,4)
y <- c(7,10,25)

fruit_out <- matrix(nrow = length(y), ncol = length(x))

for (i in seq_along(x)) {
  for (j in seq_along(y)) {
    total_cost <- y[j] * x[i]
    fruit_out[j, i] <- total_cost
  }
}

fruit_out
```


#### Try your own

```{r}

moods = c("crazy","cool","happy")
characters = c("teacher","student","surfer","roomie")

palette <- matrix(nrow = length(characters), ncol = length(moods))

for (i in seq_along(characters)) {
  for (j in seq_along(moods)) {
    mix <- paste(characters[i] , moods[j])
    palette[i,j] <- mix
  }
}
palette

```

### Logistic Growth

#### Modelng a population over time and varying growth rate

```{r}
# crate a function with a shortcut: MAC cmd + option + x

function_example <- function(x, ab) {
  x + 100 * ab
}
```


### Create the Logistic Growth Function
```{r}
logistic_growth <- function(k, n0, r, time) {
  k / (1 + ((k -n0) / n0) * exp (-r * time))
}

# Always test your functions, TEST IT!
logistic_growth(1,2,3,4)
```

```{r}
time_vec <- seq(0,35,by = 0.1)
head(time_vec)

pop_35 <- logistic_growth(n0 = 100, k = 8000, r = 0.5, time = time_vec)

pop_time35_df <- data.frame(time_vec, pop_35)

pop_time35_df$new <- c(1:length(pop_35))

```
### Graph that table

geom_jitter(aes(color = daily_air_temp_mean_c), show.legend = FALSE) +
  scale_color_gradient(low = "blue", high = "orange")
```{r}
ggplot(data = pop_time35_df, aes(x=time_vec, y= pop_35)) + 
  geom_jitter(aes(color= new)) +
  scale_color_gradient(low="blue", high = "orange")

```

```{r}
pop_35_vec <- vector(mode = "numeric", length = length(time_vec))


for (i in seq_along(time_vec)) {
  population <- logistic_growth(n0=100, k = 6000, r = 0.27, time = time_vec[i])
  pop_35_vec[i] <- population
}

```


### Now extend this to multiple growth rates

```{r}
r_seq <- seq(from = 0.2, to = 0.4, by= 0.01)
class(r_seq)
```


```{r}

out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(n0 = 100, k = 6000, r=r_seq[i], time = time_vec[j])
    out_matrix[j,i] <- pop
  }
}
```

### Now some wrangling

```{r}
logistic_df <- data.frame(out_matrix, time = time_vec)
```

### update the column names

```{r}
colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")
```

### Create pivot longer tables and data frames

```{r}
logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population",names_to = "growth_rate")


ggplot(logistic_df_long, aes(x=time, y = population)) + 
  geom_line(aes(color = growth_rate))

```


### Add a function to have error messages

Use `warning` to add a warning
Use `alert` to add a alert

$$p = 0.3\rho*A*v^3$$
rho: air density (kg / m^3)
A: area covered by turbine (square meters)
v: wind velocity (m/s)

```{r}
calc_windpower <- function(rho, radius, windspeed) {
  
  if (windspeed > 130) {
    print("FAST")
  }
  else if (rho > 1.225) {
    warning("That air density is supicious.")
  }
  if (radius < 0)
    stop("Rotor radius must be a positive value (meters) ")
  0.3*rho*pi*(radius^2)*(windspeed^3)
}
```

### Check your function

```{r}
calc_windpower(1,2,3)
```


